MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-06-07 MAESTROeX_I/algorithm.tex
readthrough of sec 3

The overall flow of the algorithm begins with a second-order Strang splitting approach to integrate the advection-reaction system for 
the thermodynamic variables $(\rho X_k, \rho h)$, followed by a second-order projection methodology to integrate the velocities subject to a divergence constraint.  Within the thermodynamic variable update we use a predictor-corrector approach to achieve second-order accuracy in time.
To summarize:
\begin{itemize}
\item In {\bf Step 1} we react the thermodynamic variables over $\Delta t/2$.
\item In {\bf Steps 2--4} we advect the thermodynamic variables over $\Delta t$.  Specifically, we compute an estimate for the expansion term, $S$, compute face-centered, time-centered velocities that satisfy the divergence constraint, and then advect the thermodynamic variables.
\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$\footnote{After this step we could skip to the velocity advance in {\bf Steps 10--11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}.
\item In {\bf Steps 6--8} we redo the advection in {\bf Steps 2--4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
\item In {\bf Step 9} we redo the reactions from {\bf Step 5} beginning with the improved results from {\bf Steps 6--8}.
\item In {\bf Steps 10--11} we advect the velocity, and then project these velocities so they satisfy the divergence constraint.
\end{itemize}

@@

The temporal integration scheme contains the following steps:
\begin{description}

%--------------------------------------------------------------------------
% STEP 1
%--------------------------------------------------------------------------
\item[Step 1] {\em React the full state through the first time interval of $\dt / 2.$}

@@

\item[Step 2] {\em Compute the provisional time-centered expansion,
    $S^{\nph,\star}$.}

@@

%--------------------------------------------------------------------------
% STEP 3
%--------------------------------------------------------------------------
\item[Step 3] {\em Construct the provisional time-centered advective velocity on
edges, $\uadvone$.}

The construction of face-centered time-centered states used to discretize the
advection terms for velocity, species, and enthalpy, are performed using
a standard multidimensional corner transport upwind approach
\citep{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
Godunov advection approach for all steps in this algorithm are described 
in Appendix A of \cite{XRB_III}.
Here we use equation (\ref{eq:momentum}) to compute time-centered edge velocities, $\uadvonedag$.
The $\dagger$ superscript refers to the fact that the predicted velocity field does not satisfy the divergence constraint,
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
\end{equation}
 We project $\uadvonedag$ onto the space of velocities that satisfy the constraint to obtain $\uadvone$.
Each projection step in the algorithm involves the solution of a variable-coefficient Poisson solve using multigrid.
The details of this ``MAC'' projection are provided in Appendix \ref{Sec:Projection}.

%--------------------------------------------------------------------------
% STEP 4
%--------------------------------------------------------------------------
\item[Step 4] {\em Advect the full state through a time interval of $\dt.$}

@@

\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  As described in Paper V, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to edges
  and here we spatially interpolate $\rho_0^n$ to edges to assemble the fluxes.

@@

\item Update the enthalpy using a discretized version of equation
%
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,
\end{equation}
%
again omitting the reaction terms since we already accounted for
them in {\bf React State}.  This equation takes the form:
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
\end{equation}
For spherical geometry, we solve the analytically equivalent form,
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
\end{equation}
The update consists of two steps:

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
  for the conservative update of $(\rho h)^{(1)}$ using using a Godunov approach \citep{XRB_III}.
  As described in Paper V, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to edges,
  where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0]$,
  and here we spatially interpolate $(\rho h)_0^n$ to edges to assemble the fluxes.

@@

%--------------------------------------------------------------------------
% STEP 5
%--------------------------------------------------------------------------
\item[Step 5] {\em React the full state through a second time interval of $\dt / 2.$}

@@

%--------------------------------------------------------------------------
% STEP 6
%--------------------------------------------------------------------------
\item[Step 6] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}

@@

%--------------------------------------------------------------------------
% STEP 7
%--------------------------------------------------------------------------
\item[Step 7] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}

@@

%--------------------------------------------------------------------------
% STEP 8
%--------------------------------------------------------------------------
\item[Step 8] {\em Advect the base state and full state through a time interval of $\dt.$}

@@

%--------------------------------------------------------------------------
% STEP 9
%--------------------------------------------------------------------------
\item[Step 9] {\em React the full state through a second time interval of $\dt / 2.$}

@@

%--------------------------------------------------------------------------
% STEP 10
%--------------------------------------------------------------------------
\item[Step 10] {\em Define the new time expansion, $S^{n+1}$.}

>>>>

The overall flow of the algorithm begins with a second-order Strang splitting approach to integrate the advection-reaction system for 
the thermodynamic variables $(\rho X_k, \rho h)$, followed by a second-order projection methodology to integrate the velocities subject to a divergence constraint.  Within the thermodynamic variable update we use a predictor-corrector approach to achieve second-order accuracy in time.
To summarize:
\begin{itemize}
\item In {\bf Step 1} we react the thermodynamic variables over the first $\Delta t/2$ interval.
\item In {\bf Steps 2--4} we advect the thermodynamic variables over $\Delta t$.  Specifically, we compute an estimate for the expansion term, $S$, compute face-centered, time-centered velocities that satisfy the divergence constraint, and then advect the thermodynamic variables.
\item In {\bf Step 5} we react the thermodynamic variables over the second $\Delta t/2$ interval\footnote{After this step we could skip to the velocity advance in {\bf Steps 10--11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}.
\item In {\bf Steps 6--8} we redo the advection in {\bf Steps 2--4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
\item In {\bf Step 9} we redo the reactions from {\bf Step 5} beginning with the improved results from {\bf Steps 6--8}.
\item In {\bf Steps 10--11} we advect the velocity, and then project these velocities so they satisfy the divergence constraint.
\end{itemize}

@@

Beginning with $(\Ub,\rho X_k,\rho h)^n$, $\pi^{n-\myhalf}$, and $(\rho_0,p_0)^n$,
the temporal integration scheme contains the following steps:
\begin{description}

%--------------------------------------------------------------------------
% STEP 1
%--------------------------------------------------------------------------
\item[Step 1] {\em React the thermodynamic variables over the first $\Delta t / 2$ interval.}

@@

\item[Step 2] {\em Compute the time-centered expansion term, $S^{\nph,\star}$.}

@@

%--------------------------------------------------------------------------
% STEP 3
%--------------------------------------------------------------------------
\item[Step 3] {\em Construct a time-centered advective velocity on edges, $\uadvone$.}

The construction of face-centered time-centered states used to discretize the
advection terms for velocity, species, and enthalpy, are performed using
a standard multidimensional corner transport upwind approach
\citep{colella1990multidimensional,saltzman1994unsplit} with the piecewise-parabolic method (PPM)
one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
Godunov advection approach for all steps in this algorithm are described 
in Appendix A of \cite{XRB_III}.
Here we use equation (\ref{eq:momentum}) to compute time-centered edge velocities, $\uadvonedag$.
The $\dagger$ superscript refers to the fact that the predicted velocity field does not satisfy the divergence constraint,
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
\end{equation}
 We project $\uadvonedag$ onto the space of velocities that satisfy the constraint to obtain $\uadvone$.
Each projection step in the algorithm involves the solution of a variable-coefficient Poisson solve using multigrid.
The details of this ``MAC'' projection are provided in Appendix \ref{Sec:Projection}.

%--------------------------------------------------------------------------
% STEP 4
%--------------------------------------------------------------------------
\item[Step 4] {\em Advect the thermodynamic variables over a time interval of $\dt.$}

@@

\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  As described in Paper V, for robust numerical slope limiting we predict 
  $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to edges
  and here we spatially interpolate $\rho_0^n$ to edges to assemble the fluxes.

@@

\item Update the enthalpy using a discretized version of equation
%
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,
\end{equation}
%
again omitting the reaction terms since we already accounted for
them in {\bf React State}.  This equation takes the form:
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\rho h\Ub) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
\end{equation}
For spherical geometry, we solve the analytically equivalent form,
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\rho h\Ub) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
\end{equation}
The update consists of two steps:

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
  for the conservative update of $(\rho h)^{(1)}$ using using a Godunov approach \citep{XRB_III}.
  As described in Paper V, for robust numerical slope limiting 
  we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to edges,
  where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0]$,
  and here we spatially interpolate $(\rho h)_0^n$ to edges to assemble the fluxes.

@@

%--------------------------------------------------------------------------
% STEP 5
%--------------------------------------------------------------------------
\item[Step 5] {\em React the thermodynamic variables over the second $\Delta t / 2$ interval.}

@@

%--------------------------------------------------------------------------
% STEP 6
%--------------------------------------------------------------------------
\item[Step 6] {\em Compute the time-centered expansion term, $S^{\nph,\star}$.}

@@

%--------------------------------------------------------------------------
% STEP 7
%--------------------------------------------------------------------------
\item[Step 7] {\em Construct a time-centered advective velocity on edges, $\uadvtwo$.}

@@

%--------------------------------------------------------------------------
% STEP 8
%--------------------------------------------------------------------------
\item[Step 8] {\em Advect the thermodynamic variables over a time interval of $\dt.$}

@@

%--------------------------------------------------------------------------
% STEP 9
%--------------------------------------------------------------------------
\item[Step 9] {\em React the thermodynamic variables over the second $\Delta t / 2$ interval.}

@@

%--------------------------------------------------------------------------
% STEP 10
%--------------------------------------------------------------------------
\item[Step 10] {\em Define the new-time expansion term, $S^{n+1}$.}
